---
title: "tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introdction  
This package integrates the basic elements in the linear regression based on BIOSTAT650. The main function *mylm* will return a list of 5 items, They are:  
*results*: dataframe, with similar format with summary(lm)  
*residuals*: The residuals of the model  
*R.square*: The R^2 which demonstrate the percentage of variance in Y explained by the model  
*R.square.adj.*: The adjusted of R^2  
*SSE*: The sum of square errors in the model  
Also, the function will automatically perform
Morgenstern, Lewis B., et al. "Fatalism, optimism, spirituality, depressive symptoms, and stroke outcome: a population-based analysis." Stroke 42.12 (2011): 3518-3523.
```{r setup}
library(qwzLinear)

```
## Testhats
```{r}
#library(testthat)
#test_check("qwzLinear", path = 'D:/zl/Umich/course/biostat625/qwzLinear/tests/')
```

##Basic Tests
```{r}
```

##Comparsion
```{r}
```
